---
title: "Hands-on Bayesianism"
author: "Andrej Arpáš"
date: "7/7/2022"
output:
  html_document:
    code_folding: show
    fig_caption: true
    highlight: tango
    theme: yeti
    toc: true
    toc_depth: 3
    toc_float: true
subtitle: "Frequentist VS Bayesian Regression"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.align = "center",
  fig.width = 7,
  fig.height = 4.2,
  dpi = 150,
  cache = TRUE
)

library(tidyverse)
library(rstan)
library(rstanarm)
library(bayesplot)
library(loo)

options(mc.cores = max(1L, parallel::detectCores() - 1L))
rstan_options(auto_write = TRUE)
set.seed(42)
color_scheme_set("blue")
```

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}
```

# Load the data

The data and some of the examples are derivatives of the sources noted in `data/README.md` (plus a couple more).

What we've got as the dependent variable is the percentage change in GOP support from 2012 to 2016.

```{r load-data}
d <- readr::read_csv("data/election_counties.csv", show_col_types = FALSE)
glimpse(d)
```

The independent variables (all continuous): `pop_change`, `age_65_plus`, `black`, `hispanic`, `hs_grad`, `undergrad`, `homeownership_rate`, `median_home_value`, `median_income`, `poverty_rate`.

# Frequentist `lm()`

You could think about what model to use, but as an exemplar we are more interested in comparing notes than in the actual analysis. We'll therefore proceed to `lm()`.

```{r lm-freq}
lm_freq <- lm(gop_support_change ~ ., data = d)
summary(lm_freq)
```

On this fit, a one-percentage-point higher Hispanic population share is associated with about $-0.15$ points of GOP-support change (holding other covariates fixed), with a conventional $p$-value far below $0.001$. The Bayesian sections ask what a full posterior says about the same slope.

# Bayesian methods

## `rstanarm::stan_glm()`

Useful references: [stan_glm](https://mc-stan.org/rstanarm/reference/stan_glm.html), [continuous outcomes](https://mc-stan.org/rstanarm/articles/continuous.html), [bayesplot MCMC](https://mc-stan.org/bayesplot/reference/MCMC-intervals.html), [Stan warnings](https://mc-stan.org/misc/warnings.html).

`rstanarm` emulates familiar R model formulas but samples with Stan. Current default priors are autoscaled Student-$t$ / normal families (not the older fixed `normal(0, 10)` / `normal(0, 5)` values from early docs). Leaving priors unspecified is still a valid weakly informative choice.

```{r stan-glm, results='hide'}
post1 <- stan_glm(
  gop_support_change ~ .,
  data = d,
  family = gaussian(),
  seed = 42,
  refresh = 0
)
```

```{r stan-glm-summary}
print(post1, digits = 3)
loo1 <- loo(post1)
loo1

diag1 <- tibble(
  max_rhat = max(rhat(post1), na.rm = TRUE),
  min_neff_ratio = min(neff_ratio(post1), na.rm = TRUE)
)
diag1
```

```{r stan-glm-line}
draws <- as.data.frame(as.matrix(post1))
colnames(draws)[1] <- "intercept"

ggplot(d, aes(x = hispanic, y = gop_support_change)) +
  geom_point(size = 0.3, color = "#003366", alpha = 0.35) +
  geom_abline(
    data = draws, aes(intercept = intercept, slope = hispanic),
    color = "skyblue", linewidth = 0.2, alpha = 0.05
  ) +
  geom_abline(
    intercept = coef(post1)["(Intercept)"],
    slope = coef(post1)["hispanic"],
    color = "skyblue4", linewidth = 1
  ) +
  theme_minimal()
```

```{r stan-glm-ppc}
pp_check(post1, plotfun = "hist", nreps = 5)
```

`mcmc_intervals()` defaults to 50% (thick) and 90% (thin) central intervals with the point at the posterior median.

```{r stan-glm-intervals}
slope_pars <- setdiff(names(coef(post1)), c("(Intercept)"))
mcmc_intervals(
  as.matrix(post1),
  pars = slope_pars,
  prob = 0.5,
  prob_outer = 0.9
)
```

```{r stan-glm-quantiles}
m1 <- as.matrix(post1)
median(m1[, "hispanic"])
round(quantile(m1[, "hispanic"], probs = c(0.1, 0.5, 0.9)), 3)
```

```{r stan-glm-intervals-wide}
mcmc_intervals(
  as.matrix(post1),
  pars = slope_pars,
  prob = 0.8,
  prob_outer = 0.99
)
```

Conditional on the data and this model, the posterior for the Hispanic slope is tightly concentrated near $-0.15$: a 99% central interval is roughly $-0.17$ to $-0.13$.

```{r stan-glm-areas}
mcmc_areas(as.matrix(post1), pars = c("age_65_plus", "undergrad", "poverty_rate"))
```

Age $65+$ and poverty rate pull toward positive GOP-support change in this period; bachelor’s share pulls negative, with a narrower central mass.

```{r stan-glm-pp-dens}
pp_check(post1)
```

The density overlay shows mild mismatch in the bulk (medians slightly off), which could be remedied e.g. by changing our priors.

## Custom Stan (`rstan`)

Writing the likelihood and priors in Stan makes the model explicit. Below is one centered/scaled design with Student-$t$ priors (`stan/mlr.stan`). Flat-prior versions of this regression mix poorly on modern Stan; weakly informative priors are the pedagogical default here.

```{r stan-data}
X_scaled <- scale(as.matrix(d |> select(-gop_support_change)))
stan_dat <- list(
  n = nrow(d),
  k = ncol(X_scaled),
  X = X_scaled,
  Y = d$gop_support_change
)
```

```{r stan-fit, results='hide'}
post2 <- stan(
  file = "stan/mlr.stan",
  data = stan_dat,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 42,
  refresh = 0,
  control = list(adapt_delta = 0.95)
)
```

```{r stan-fit-diag}
sum2 <- summary(post2)$summary
diag2 <- tibble(
  max_rhat = max(sum2[, "Rhat"], na.rm = TRUE),
  min_n_eff = min(sum2[, "n_eff"], na.rm = TRUE)
)
diag2

beta_pars <- paste0("beta[", seq_len(stan_dat$k), "]")
color_scheme_set("yellow")
mcmc_intervals(post2, pars = beta_pars) +
  scale_y_discrete(labels = colnames(X_scaled))
```

```{r stan-fit-areas}
color_scheme_set("yellow")
mcmc_areas(post2, pars = c("beta[4]", "beta[7]")) +
  labs(title = "Posterior Distributions") +
  scale_y_discrete(labels = c("hispanic", "homeownership_rate"))
color_scheme_set("blue")
```

Slopes here are in **scaled-predictor units**, so magnitudes are not directly comparable to the `stan_glm` coefficients on the raw scale. The qualitative signs for Hispanic and homeownership still line up.

# Prior elicitation with GLD / J-QPD

Johnson quantile-parameterized distributions (J-QPD) and the generalized lambda distribution (GLD) let you state priors through quantiles rather than a named family. Helpers live in `stan/quantile_functions.stan` and `R/GLD_helpers.R`.

Draw $S$ prior predictive $R^2$ values:

$$
R_s^2=\frac{\frac{1}{N-1}\sum_{n=1}^{N}(\tilde\mu_n-\bar\mu)^2}{\tilde\sigma^2+\frac{1}{N-1}\sum_{n=1}^{N}(\tilde\mu_n-\bar\mu)^2}.
$$

```{r expose-gld}
rstan::expose_stan_functions("stan/quantile_functions.stan")
source("R/GLD_helpers.R")

X_c <- model.matrix(gop_support_change ~ ., data = d)[, -1, drop = FALSE]
X_c <- sweep(X_c, 2, colMeans(X_c), `-`)
```

```{r gld-elicit}
m_alpha <- -10.9
s_alpha <- 4.1

lowers <- c(
  PC = -0.0005, A = 0.0001, B = -0.0004, H = -0.0004, HS = 0.0023,
  U = -0.0009, HR = -0.0003, MHV = -0.0003, MI = -0.0003, PR = -0.0004
)
medians <- c(
  PC = -0.0004, A = 0.00015, B = -0.0003, H = -0.0003, HS = 0.0025,
  U = -0.0008, HR = -0.0002, MHV = -0.0002, MI = -0.0002, PR = -0.0003
)
uppers <- c(
  PC = -0.0003, A = 0.0002, B = -0.0002, H = -0.0002, HS = 0.0027,
  U = -0.0007, HR = -0.0001, MHV = -0.0001, MI = -0.0001, PR = -0.0002
)
ninety <- c(
  PC = -0.0002, A = 0.00025, B = -0.0001, H = -0.0001, HS = 0.0029,
  U = -0.0006, HR = 0.00, MHV = 0.00, MI = 0.00, PR = -0.0001
)

a_s <- lapply(seq_along(medians), function(i) {
  GLD_solver(lowers[i], medians[i], uppers[i], ninety[i], alpha = 0.9)
})
names(a_s) <- names(medians)

q_sigma <- c(lower = 0.25, median = 0.5, upper = 1)
```

These quantile targets are deliberately tight and inspired by the `stan_glm` posterior from the preceding chunks.

```{r prior-r2}
feat <- c("pop_change", "age_65_plus", "black", "hispanic", "hs_grad",
          "undergrad", "homeownership_rate", "median_home_value",
          "median_income", "poverty_rate")
keys <- c("PC", "A", "B", "H", "HS", "U", "HR", "MHV", "MI", "PR")

R_squared <- replicate(1000, {
  alpha_ <- rnorm(1, mean = m_alpha, sd = s_alpha)
  betas <- vapply(seq_along(keys), function(j) {
    k <- keys[j]
    GLD_rng(
      medians[k],
      IQR = uppers[k] - lowers[k],
      asymmetry = a_s[[k]][1],
      steepness = a_s[[k]][2]
    )
  }, numeric(1))
  mu_ <- as.numeric(alpha_ + X_c[, feat, drop = FALSE] %*% betas)
  sigma_ <- JQPDS_rng(lower_bound = 0, alpha = 0.25, quantiles = q_sigma)
  numer <- var(mu_)
  numer / (sigma_^2 + numer)
})
summary(R_squared)
```

## Posterior with an $R^2$ prior (`stan_lm`)

```{r stan-lm, results='hide'}
post4 <- stan_lm(
  gop_support_change ~ .,
  data = d,
  prior = R2(location = median(R_squared), what = "median"),
  prior_intercept = normal(location = m_alpha, scale = s_alpha, autoscale = FALSE),
  seed = 42,
  refresh = 0
)
```

```{r stan-lm-summary}
print(post4, digits = 2)
diag4 <- tibble(
  max_rhat = max(rhat(post4), na.rm = TRUE),
  min_neff_ratio = min(neff_ratio(post4), na.rm = TRUE)
)
diag4
loo4 <- loo(post4)
loo_compare(loo1, loo4)
```

```{r stan-lm-ppc}
ppc_intervals(
  y = d$gop_support_change,
  yrep = posterior_predict(post4),
  x = d$hispanic
) +
  xlab("hispanic")
```

```{r stan-lm-table}
m4 <- as.matrix(post4)
slope_names <- setdiff(colnames(m4), c("(Intercept)", "sigma", "R2", "log-fit_ratio"))
slope_names <- slope_names[slope_names %in% colnames(d)]

posterior_quantiles <- purrr::map_dfr(slope_names, function(nm) {
  q <- quantile(m4[, nm], probs = c(0.025, 0.25, 0.5, 0.75, 0.975))
  tibble(
    term = nm,
    q025 = unname(q[1]),
    q25 = unname(q[2]),
    q50 = unname(q[3]),
    q75 = unname(q[4]),
    q975 = unname(q[5])
  )
})
posterior_quantiles
```

# Conclusion

Three fits on the same county panel:

1. **OLS** — point estimates and $p$-values.
2. **`stan_glm`** — default weakly informative priors, full posterior, LOO, PPC.
3. **Custom Stan + `stan_lm($R^2$)`** — explicit Student-$t$ priors on a scaled design, then a quantile-elicited $R^2$ prior that folds GLD / J-QPD into `rstanarm`.

It may not be quite trivial to "guess" the prior stuff using quantile functions. But the tradeoff is that we can dispense with pre-defined distributions — a cleverer approach, but more difficult to implement.
