---
title: "CDSS Hackathon 2020"
author: "Korina Y. Baraceros & Andrej Arpáš"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    fig_caption: true
    theme: spacelab
    code_folding: show
subtitle: "DK & SE — Time Series on Covid-19 Data Using Stan"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.align = "center",
  dpi = 120
)

library(tidyverse)
library(naniar)
library(lubridate)
library(gridExtra)
library(rstanarm)
library(bayesplot)

options(scipen = 9999)
options(mc.cores = max(1L, parallel::detectCores() - 1L))
set.seed(42)
color_scheme_set("teal")

col_dnk <- "#14b1ab"
col_swe <- "#f0a500"
col_close <- "#de4463"
```

## Background

How has COVID-19 affected mobility of people and goods across the globe?

We are interested in the pandemic's effects on society. With so many different and disparate approaches, it is difficult to see what measures correlated with what behavior. We decided to compare two regions, similar in geographic location, population size, culture, and history: **Sweden** and **Denmark**. Their mobility behavior and COVID-19 statistics are reflected in Google Community Mobility Reports and ECDC case/death series (see `data/README.md`).

We chose these regions because the approach to mitigating the effects of COVID-19 in terms of mobility and the resultant economic and health implications were considerably different. While both effectively closed their borders, mobility was severely curtailed in Denmark (lockdown) whereas Sweden only issued recommendations (non-mandatory). These are contrasting responses for geographies that are physically connected and whose populations are virtually homogeneous to each other.

An important consideration for these datasets is to treat time itself as a variable. Here we consider time in terms of **before-and-after** border closures, and we employ a before-and-after modeling technique called **difference-in-differences** (pretest-posttest with comparison group). For this to be a valid method, we assume parallel trends — reflected in our choice of Sweden and Denmark.

To incorporate uncertainty we use Bayesian modeling via Stan in R (`rstanarm::stan_glm`), which gives a posterior distribution for our estimates rather than single point estimates.

```{r load-clean}
mobility_raw <- read_csv("data/google_mobility_swe_den.csv", show_col_types = FALSE)
covid_raw <- read_csv("data/covid19_ecdc.csv", show_col_types = FALSE)

covid19 <- covid_raw |>
  filter(countries_and_territories %in% c("Sweden", "Denmark"), year == 2020) |>
  arrange(date) |>
  transmute(
    country = countries_and_territories,
    date,
    daily_deaths,
    daily_confirmed_cases,
    deaths,
    pop = pop_data_2019,
    daily_deaths_per_mil = 1e6 * daily_deaths / pop_data_2019,
    daily_confirmed_cases_per_mil = 1e6 * daily_confirmed_cases / pop_data_2019,
    deaths_per_mil = 1e6 * deaths / pop_data_2019
  )

mobility_region <- mobility_raw |>
  filter(country_region %in% c("Sweden", "Denmark")) |>
  transmute(
    country = country_region,
    region = sub_region_1,
    date,
    retail_rec = retail_and_recreation_percent_change_from_baseline,
    grocery_pharm = grocery_and_pharmacy_percent_change_from_baseline,
    parks = parks_percent_change_from_baseline,
    transit = transit_stations_percent_change_from_baseline,
    workplace = workplaces_percent_change_from_baseline,
    residential = residential_percent_change_from_baseline
  ) |>
  mutate(
    yr_month = format(as.Date(date), "%Y-%m"),
    week = week(ymd(date))
  )

# Country-day mobility means (avoid broadcasting country COVID onto every region).
mobility_cd <- mobility_region |>
  group_by(country, date, yr_month, week) |>
  summarise(
    across(
      c(retail_rec, grocery_pharm, parks, transit, workplace, residential),
      ~ mean(.x, na.rm = TRUE)
    ),
    .groups = "drop"
  )

panel <- mobility_cd |>
  left_join(covid19, by = c("country", "date")) |>
  mutate(
    # Shared post period: after Sweden's non-EU border cutoff (DK closed Mar 14; SE Mar 19).
    closed = as.integer(date > as.Date("2020-03-19")),
    country = factor(country, levels = c("Denmark", "Sweden"))
  )

tibble(
  n_region_day = nrow(mobility_region),
  n_country_day = nrow(panel),
  date_min = min(panel$date),
  date_max = max(panel$date)
)
```

Time series models are sensitive to missingness, so check the region-day join (before the country-day collapse used in the DiD).

```{r missingness, fig.height=11, fig.width=13}
mobility_vis <- mobility_region |>
  left_join(covid19, by = c("country", "date"))

# Keep column order; put labels underneath with room so long names are not clipped.
p_miss <- vis_miss(mobility_vis, warn_large_data = FALSE)
x_levels <- levels(p_miss$data$variable)
p_miss +
  scale_x_discrete(limits = x_levels, position = "bottom") +
  theme(
    axis.text.x = element_text(angle = 55, hjust = 1, vjust = 1, size = 8.5),
    plot.margin = unit(c(0.5, 1.0, 4.0, 0.5), "cm")
  )
```

The data seems to have been collected sparsely for some variables on some days, so we will aggregate into weeks for the charts below, and the DiD models use the country-day panel. The jitter from day to day is likely too granular for this model anyway.

## Data exploration

### After border closures

Denmark closed its border on **March 14, 2020**. Sweden closed its borders to non-EU member states on **March 19, 2020**, and virtually all EU countries imposed travel restrictions at about the same time, effectively prohibiting travel into Sweden. Both are closed within a week of each other; we treat **2020-03-19** as the shared post-period cutoff.

### Change in mobility

```{r eda-mobility-week, fig.height=12, fig.width=10}
mob_week <- panel |>
  group_by(country, week) |>
  summarise(
    across(
      c(retail_rec, grocery_pharm, transit, workplace, residential),
      ~ mean(.x, na.rm = TRUE)
    ),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = -c(country, week),
    names_to = "series",
    values_to = "mean"
  ) |>
  mutate(
    series = recode(
      series,
      retail_rec = "Retail",
      grocery_pharm = "Grocery & Pharmacy",
      transit = "Transit",
      workplace = "Workplace",
      residential = "Residential"
    )
  )

ggplot(mob_week, aes(week, mean, colour = country, group = country)) +
  geom_point(size = 1.2) +
  geom_line(linewidth = 0.6) +
  geom_vline(xintercept = 12, linetype = "dotted", colour = col_close, linewidth = 0.85) +
  facet_wrap(~series, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c(Denmark = col_dnk, Sweden = col_swe)) +
  labs(
    title = "Changes in mobility over time by week",
    x = "Week",
    y = "% change from baseline",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

### Change in COVID metrics

```{r eda-covid-week, fig.height=8, fig.width=10}
covid_week <- panel |>
  group_by(country, week) |>
  summarise(
    daily_confirmed_cases_per_mil = mean(daily_confirmed_cases_per_mil, na.rm = TRUE),
    daily_deaths_per_mil = mean(daily_deaths_per_mil, na.rm = TRUE),
    deaths_per_mil = mean(deaths_per_mil, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = -c(country, week),
    names_to = "series",
    values_to = "mean"
  ) |>
  mutate(
    series = recode(
      series,
      daily_confirmed_cases_per_mil = "Daily confirmed cases per million",
      daily_deaths_per_mil = "Daily deaths per million",
      deaths_per_mil = "Cumulative deaths per million"
    )
  )

ggplot(covid_week, aes(week, mean, colour = country, group = country)) +
  geom_point(size = 1.2) +
  geom_line(linewidth = 0.6) +
  geom_vline(xintercept = 12, linetype = "dotted", colour = col_close, linewidth = 0.85) +
  facet_wrap(~series, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c(Denmark = col_dnk, Sweden = col_swe)) +
  labs(
    title = "Weekly averages of COVID-19 statistics",
    x = "Week",
    y = NULL,
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

## Difference in differences

**What is the effect of a mobility-related policy on COVID-related metrics?**

A difference-in-differences model compares COVID metrics before-and-after border closure for Sweden to the before-and-after for Denmark. For two countries that can be considered quite easily the same country, we would expect the closure of borders to have comparable effects on COVID metrics under parallel trends. The DiD interaction `closed:countrySweden` is the quantity of interest.

### In conditional mean terms

```{r did-means}
did_means <- function(y) {
  panel |>
    filter(!is.na(.data[[y]])) |>
    group_by(country, closed) |>
    summarise(mean = mean(.data[[y]], na.rm = TRUE), .groups = "drop") |>
    mutate(outcome = y)
}

means_cases <- did_means("daily_confirmed_cases_per_mil")
means_deaths <- did_means("daily_deaths_per_mil")

cell <- function(tbl, ctry, cl) {
  tbl$mean[tbl$country == ctry & tbl$closed == cl]
}

delta <- function(tbl, ctry) cell(tbl, ctry, 1L) - cell(tbl, ctry, 0L)

tibble(
  outcome = c("daily_confirmed_cases_per_mil", "daily_deaths_per_mil"),
  sweden_before_after = c(delta(means_cases, "Sweden"), delta(means_deaths, "Sweden")),
  denmark_before_after = c(delta(means_cases, "Denmark"), delta(means_deaths, "Denmark")),
  naive_did = c(
    delta(means_cases, "Sweden") - delta(means_cases, "Denmark"),
    delta(means_deaths, "Sweden") - delta(means_deaths, "Denmark")
  )
)
```

Observing only differences in means of daily confirmed cases per million, Sweden's before→after change exceeds Denmark's by the naive DiD above (same construction for daily deaths per million).

### In regression terms

Primary specification on the country-day panel:

\[
y_{ct} = \alpha + \beta_1\,\mathrm{closed}_t + \beta_2\,\mathrm{Sweden}_c + \beta_3\,(\mathrm{closed}_t\times\mathrm{Sweden}_c) + \varepsilon_{ct}.
\]

$\beta_3$ is the DiD estimate. A second model adds retail and grocery/pharmacy mobility as covariates.

```{r did-cases, results='hide'}
reg_cases <- stan_glm(
  daily_confirmed_cases_per_mil ~ closed * country,
  data = panel,
  family = gaussian(),
  seed = 42,
  refresh = 0
)

reg_cases_mob <- stan_glm(
  daily_confirmed_cases_per_mil ~ closed * country + retail_rec * grocery_pharm,
  data = panel,
  family = gaussian(),
  seed = 42,
  refresh = 0
)
```

```{r did-cases-summary}
print(reg_cases, digits = 3)
print(reg_cases_mob, digits = 3)

diag_cases <- tibble(
  model = c("DiD", "DiD + mobility"),
  max_rhat = c(max(rhat(reg_cases), na.rm = TRUE), max(rhat(reg_cases_mob), na.rm = TRUE)),
  min_neff_ratio = c(
    min(neff_ratio(reg_cases), na.rm = TRUE),
    min(neff_ratio(reg_cases_mob), na.rm = TRUE)
  )
)
diag_cases

posterior_interval(reg_cases, prob = 0.9)
posterior_interval(reg_cases_mob, prob = 0.9)
```

```{r did-cases-ppc, fig.height=3.8}
pp_check(reg_cases) + ggtitle("PPC: daily cases per million (DiD)")
```

```{r did-deaths, results='hide'}
reg_deaths <- stan_glm(
  daily_deaths_per_mil ~ closed * country,
  data = panel,
  family = gaussian(),
  seed = 42,
  refresh = 0
)

reg_deaths_mob <- stan_glm(
  daily_deaths_per_mil ~ closed * country + retail_rec * grocery_pharm,
  data = panel,
  family = gaussian(),
  seed = 42,
  refresh = 0
)
```

```{r did-deaths-summary}
print(reg_deaths, digits = 3)
print(reg_deaths_mob, digits = 3)

diag_deaths <- tibble(
  model = c("DiD", "DiD + mobility"),
  max_rhat = c(max(rhat(reg_deaths), na.rm = TRUE), max(rhat(reg_deaths_mob), na.rm = TRUE)),
  min_neff_ratio = c(
    min(neff_ratio(reg_deaths), na.rm = TRUE),
    min(neff_ratio(reg_deaths_mob), na.rm = TRUE)
  )
)
diag_deaths

posterior_interval(reg_deaths, prob = 0.9)
posterior_interval(reg_deaths_mob, prob = 0.9)
```

```{r did-deaths-ppc, fig.height=3.8}
pp_check(reg_deaths) + ggtitle("PPC: daily deaths per million (DiD)")
```

```{r did-interaction-plot, fig.height=3.6}
mcmc_areas(as.matrix(reg_cases), pars = "closed:countrySweden") +
  labs(title = "Posterior: DiD interaction (cases per million)")
mcmc_areas(as.matrix(reg_deaths), pars = "closed:countrySweden") +
  labs(title = "Posterior: DiD interaction (deaths per million)")
```

Read the interaction posteriors as Sweden's excess before→after change relative to Denmark, after the shared border-closure cutoff. Mobility covariates in the secondary models absorb some of the time-varying behavior associated with retail/grocery activity; the interaction remains the DiD contrast of interest.

## Discussion and next steps

Our results speak to whether, irrespective of mobility curtailment and measures undertaken — from border closures to restaurant shutdowns — both countries' health metrics eventually moved in ways that leave only marginal differences through time, while much of the observed contrast sits in diminished mobility and increased economic disruption. Next up, it would be prudent to examine data pertaining to the characteristics of the deceased and infected. Should additional data confirm what the reportage suggests, i.e. that the majority of the deaths occurred concomitant with underlying health issues, especially in higher age brackets, the rationale for mobility restrictions would face serious credibility problems.
